# Environment

This section discusses the usage of functions which manipulate environments.

## baseenv()

This is the environment of the base package.

```{r echo=FALSE, message=FALSE, out.width = "575px"}
show_graph("environment", "baseenv.png")
show_graph("environment", "baseenv-relative.png")
```

## emptyenv()

This is the empty environment, the topmost environment.

```{r echo=FALSE, message=FALSE, out.width = "575px"}
show_graph("environment", "emptyenv.png")
show_graph("environment", "emptyenv-relative.png")
```

## new.env(hash = TRUE, parent = parent.frame(), size = 29L)

`new.env` returns a new (empty) environment with (by default) enclosure the parent frame.

```{r echo=FALSE, message=FALSE, out.width = "575px"}
show_graph("environment", "new.env.png")
show_graph("environment", "new.env-relative.png")
```

## NewEnvironment

This is the C level function which gets called when a new environment is created.

```{r echo=FALSE, message=FALSE, out.width = "575px"}
show_graph("environment", "NewEnvironment.png")
show_graph("environment", "NewEnvironment-relative.png")
```

## environment(fun = NULL)

If fun is a function or a formula then `environment(fun)` returns the environment associated with that function or formula. If fun is NULL then the current evaluation environment is returned.

```{r echo=FALSE, message=FALSE, out.width = "575px"}
show_graph("environment", "environment.png")
show_graph("environment", "environment-relative.png")
```

## environment(fun) <- value

`environment(fun) <- value` sets the environment of the function or formula fun to the value given.

```{r echo=FALSE, message=FALSE, out.width = "575px"}
show_graph("environment", "environment<-.png")
show_graph("environment", "environment<--relative.png")
```

## parent.env

`parent.env` returns the enclosing environment of its argument.

```{r echo=FALSE, message=FALSE, out.width = "575px"}
show_graph("environment", "parent.env.png")
show_graph("environment", "parent.env-relative.png")
```

## parent.env(env) <- value

`parent.env<-` sets the enclosing environment of its first argument.
The replacement function parent.env<- is extremely dangerous as it can be used to destructively change environments in ways that violate assumptions made by the internal C code. It may be removed in the near future.

```{r echo=FALSE, message=FALSE, out.width = "575px"}
show_graph("environment", "parent.env<-.png")
show_graph("environment", "parent.env<--relative.png")
```

## Data (Quartiles)
```{r echo=FALSE, message=FALSE}
get_table("environment", "environment_function_usage.csv") %>%
select(-`SCRIPT`) %>%
group_by(`FUNCTION NAME`) %>%
summarize(`MEAN` = mean(`COUNT`),
          `MEDIAN` = median(`COUNT`),
          `LOWER QUANTILE (25%)` = unname(stats::quantile(sort(`COUNT`), 0.25)),
          `UPPER QUANTILE (75%)` = unname(stats::quantile(sort(`COUNT`), 0.75))) %>%
datatable(rownames=FALSE)
```

## Data (Summarized)
```{r echo=FALSE, message=FALSE}
get_table("environment", "environment_function_usage.csv") %>%
group_by(`FUNCTION NAME`) %>%
summarize(`COUNT` = sum(`COUNT`)) %>%
datatable(rownames = FALSE)
```

## Data (Unsummarized)

```{r echo=FALSE, message=FALSE}
show_table("environment", "environment_function_usage.csv")
```
