# Side Effects

This section describes the kind of operations performed by promises.
This gives an insight into how many promises are pure and how many
perform side effects.

## Action Count

There are four types of environment actions -

1. **asn** - assigning to a variable in an environment
2. **rea** - reading a variable from an environment
3. **def** - defining a new variable in an environment
4. **rem** - removing a variable from the environment

The graph below shows the number of times these actions are
performed by promises.

```{r echo=FALSE, message=FALSE}
show_graph("side-effects", "action_count_distribution.png")
```

### Data (Summarized)

```{r echo=FALSE, message=FALSE}
show_table("side-effects", "action_count_distribution_total.csv")
```

### Data

```{r echo=FALSE, message=FALSE}
show_table("side-effects", "action_count_distribution.csv")
```

## Promise count by action

This section shows the number of distinct promises that perform
a particular environment action.
Another category **none** gives the number of promises
that are evaluated but don't even perform a read operation.

```{r echo=FALSE, message=FALSE}
show_graph("side-effects", "action_promise_count_distribution.png")
```

### Raw Data (Summarized)

```{r echo=FALSE, message=FALSE}
show_table("side-effects", "action_promise_count_distribution_total.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("side-effects", "action_promise_count_distribution.csv")
```

## Promise count by purity

This section distinguishes between promises that perform 
side effects and promises that are pure.
Pure promises are those that only perform the **rea** 
environment action.
Benign promises are those that don't even perform the 
**rea** operation.


```{r echo=FALSE, message=FALSE}
show_graph("side-effects", "promise_purity_distribution.png")
```

### Raw Data (Summarized)

```{r echo=FALSE, message=FALSE}
show_table("side-effects", "promise_purity_distribution_total.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("side-effects", "promise_purity_distribution.csv")
```
