# Lifespan

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 1200)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(methods) ## including this prevents a warning - "executing load actions ..."
library(DT)
library(tidyverse)
library(knitr)
library(png)
library(grid)
library(gridExtra)
```

```{r echo=FALSE, message=FALSE}
# https://www.rdocumentation.org/packages/knitr/versions/1.17/topics/include_graphics

get_graph <- function(analysis_name, graph_filename) {
    file.path(params$analysis_output_dir, analysis_name, "graph", graph_filename)
}

show_graph <- function(analysis_name, graph_filename) {
    get_graph(analysis_name, graph_filename) %>%
    include_graphics()
}

get_table <- function(analysis_name, table_filename) {
    file.path(params$analysis_output_dir, analysis_name, "table", table_filename) %>%
    read_csv()
}

show_table <- function(analysis_name, table_filename) {
    get_table(analysis_name, table_filename) %>%
    datatable(rownames = FALSE)
}
```

```{r echo=FALSE, message=FALSE}
lifespan_summary <- get_table("promise-lifespan", "lifespan_summary.csv")
required_lifespan_summary <- get_table("promise-lifespan", "required_lifespan_summary.csv")
```

This section discusses the lifespan of promises in terms of the number of GC Cycles.
The lifespan of a promise is defined as the number of GC cycles for which the 
promise is alive.

## Promise Categories

We categorize the promises observed while tracing as follows:

1. **Foreign Promises** are created before the tracing starts. Since we don't see their creation, 
   we can't study their lifespan in terms of GC cycles. They are **excluded from the analysis**
2. **Immortal Promises** are created but not garbage collected while tracing. They are 
   **excluded from the analysis** because we can't compute their lifespan.
3. **Mortal Promises** are created and garbage collected while tracing.
   They are the only promises that are **included in the analysis**.
   They are further classified as follows:
   i. **Forced Promises** are the mortal promises that are looked up at least once.
   ii. **Unforced Promises** are mortal promises that are never looked up.

- **TOTAL PROMISE COUNT** = **FOREIGN PROMISE COUNT** + **IMMORTAL PROMISE COUNT** + **MORTAL PROMISE COUNT**

- **IMMORTAL PROMISE COUNT** = **UNFORCED PROMISE COUNT** + **FORCED PROMISE COUNT**

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "category_counts.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "count_summary.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "counts.csv")
```

## Mean GC Cycles Distribution

Following graph shows the weighted mean of the number GC cycles 
for which promises are (should be) alive in the scripts.

```{r echo=FALSE, message=FALSE, out.width="500px"}
show_graph("promise-lifespan", "lifespan_mean_gc_cycles.png")
show_graph("promise-lifespan", "required_lifespan_mean_gc_cycles.png")
show_graph("promise-lifespan", "extra_lifespan_mean_gc_cycles.png")
```

## Median GC Cycles Distribution

Following graph shows the median number GC cycles for which 
promises are (should be) alive in the scripts.

```{r echo=FALSE, message=FALSE, out.width="500px"}
show_graph("promise-lifespan", "lifespan_median_gc_cycles.png")
show_graph("promise-lifespan", "required_lifespan_median_gc_cycles.png")
show_graph("promise-lifespan", "extra_lifespan_median_gc_cycles.png")
```

## Min GC Cycles Distribution

Following graph shows the minimum number GC cycles for which 
promises are (should be) alive in the scripts.

```{r echo=FALSE, message=FALSE, out.width="500px"}
show_graph("promise-lifespan", "lifespan_min_gc_cycles.png")
show_graph("promise-lifespan", "required_lifespan_min_gc_cycles.png")
show_graph("promise-lifespan", "extra_lifespan_min_gc_cycles.png")
```

## Max GC Cycles Distribution

Following graph shows the maximum number GC cycles for which 
promises are (should be) alive in the scripts.

```{r echo=FALSE, message=FALSE, out.width="500px"}
show_graph("promise-lifespan", "lifespan_max_gc_cycles.png")
show_graph("promise-lifespan", "required_lifespan_max_gc_cycles.png")
show_graph("promise-lifespan", "extra_lifespan_max_gc_cycles.png")
```

## Lifespan Data (Summarized)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "lifespan_summary.csv")
```

## Required Lifespan Data (Summarized)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "required_lifespan_summary.csv")
```

## Extra Lifespan Data (Summarized)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "extra_lifespan_summary.csv")
```


## Lifespan Data (Raw)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "lifespan_summary.csv")
```

## Required Lifespan Data (Raw)
```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "required_lifespan.csv")
```

<!--
### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "alive-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "alive.csv")
```

## Indispensable Promises

The following graph shows the number of mortal promises that need to be alive for a given number of GC cycles.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "indispensable.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "indispensable-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "indispensable.csv")
```


## Dispensable Promises

The following graph shows the number of promises that extend their required lifespan by a given number of GC cycles.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "dispensable.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "dispensable-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "dispensable.csv")
```


## Unforced Promises

The following graph shows the number of promises that are never forced.

```{r echo=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
show_graph("promise-lifespan", "unforced.png")
```

### Summarized Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "unforced-summarized.csv")
```

### Raw Data

```{r echo=FALSE, message=FALSE}
show_table("promise-lifespan", "unforced.csv")
```


## Comparison

```{r echo=FALSE, message=FALSE}

indispensable <- 
    get_table("promise-lifespan", "indispensable-summarized.csv") %>%
    rename(`INDISPENSABLE PROMISE COUNT` = `PROMISE COUNT`)

dispensable <- 
    get_table("promise-lifespan", "dispensable-summarized.csv") %>%
    rename(`DISPENSABLE PROMISE COUNT` = `PROMISE COUNT`)

alive <- 
    get_table("promise-lifespan", "alive-summarized.csv") %>%
    rename(`ALIVE PROMISE COUNT` = `PROMISE COUNT`)

alive %>%
full_join(indispensable, by = c("GC CYCLES" = "GC CYCLES")) %>%
full_join(dispensable, by = c("GC CYCLES" = "GC CYCLES")) %>%
datatable(rownames = FALSE)
```
-->
