# Byte-code compilation

```{r, "byte-code compilation, calls", echo=FALSE, fig.width=5, fig.height=3.5, warning=FALSE}
    closure_call_compilation <- load_local_csv("compiled_calls.csv", tibble(compiled=logical(), number=numeric(), percent=numeric()))
    
    data <- 
      closure_call_compilation %>% 
      group_by(compiled) %>% 
      summarise(number=sum(as.numeric(number))) %>% 
      mutate(percent=100*number/sum(number))
    
    if (data %>% count %>% pull(n) > 0) {
      g<-ggplot(data, aes(x=compiled, y=number)) + 
        geom_bar(stat="identity") +
        ylab("No. of calls") + 
        scale_y_continuous(labels=pp_trunc) +
        theme(axis.title.x=element_blank())
      g
      save_graph("byte-code-compilation-calls.pdf", g)
    }
    
    if (data %>% count %>% pull(n) > 0) {
      data %>% prettify %>% kable(col.names=c("compiled","number","percentage"), align="lrr")
    }
```

```{r, "byte-code compilation, functions", echo=FALSE, fig.width=5, fig.height=3.5, warning=FALSE}
    closure_function_compilation <- load_local_csv("aggregated_compiled_functions.csv", tibble(compiled=logical(), number=numeric(), percent=numeric()))
    
    data <- 
      closure_function_compilation %>% 
      group_by(compiled) %>% 
      summarise(number=sum(as.numeric(number))) %>% 
      mutate(percent=100*number/sum(number))
    
    if (data %>% count %>% pull(n) > 0) {
      g <- ggplot(data, aes(x=compiled, y=number)) + 
        geom_bar(stat="identity") +
        ylab("No. of calls") + 
        scale_y_continuous(labels=pp_trunc) +
        theme(axis.title.x=element_blank())
      g
      save_graph("byte-code-compilation-functions.pdf", g)
    }
      
    if (data %>% count %>% pull(n) > 0) {
      data %>% prettify %>% kable(col.names=c("compiled","number","percentage"), align="lrr")
    }
```
    
```{r, "byte-code compilation, functions loading", echo=FALSE, warning=FALSE}
    #closure_function_compilation <- load_local_csv("compiled_functions.csv")
```
    
      
      <div style="font-size: 50%">
      When calling a function it's either byte-code **compiled** or **uncompiled**. When agregated across all calls to a function, a function can be:  
    **compiled** - all calls to this function are compiled,  
    **uncompiled** - none of the calls to this function are compiled,  
    **after 1st** - a function can be compiled by the JIT, so it can be uncompiled initially and then compiled later on,  
    **erratic** - a function is sometimes compiled and sometimes not; it can indicate that two separate functions are indistinguishable to the tracer (same definition, unknown location) and one of them is compiled while another is not.
    </div>